---
title: "Research in Progress"

format:
  html:
    mermaid: 
      theme: neutral
---

```{r}
#| include: false
#| echo: false

source("0_packages.R")

make_mermaid_flow <- function(current_level = "B") {
  
  levels <- c("A", "B", "C", "D", "E")
  
  labels <- c(
    A = "Formal analysis plan",
    B = "Execute analysis",
    C = "Draft manuscript",
    D = "Submit to journal",
    E = "Manuscript accepted"
  )
  
  # Determine status
  current_idx <- match(current_level, levels)
  
  status <- purrr::map_chr(seq_along(levels), \(i) {
    if (i < current_idx) {
      "done"
    } else if (i == current_idx) {
      "current"
    } else {
      NA_character_
    }
  })
  
  names(status) <- levels
  
  # Build nodes + edges
  edges <- paste0(
    levels[-length(levels)], "[/", labels[levels[-length(levels)]], "/] ~~~ ",
    levels[-1], "[/", labels[levels[-1]], "/]"
  )
  
  # Build class statements
  class_lines <- purrr::imap_chr(status, \(s, lvl) {
    if (is.na(s)) return(NA_character_)
    paste0("class ", lvl, " ", s, ";")
  }) |> stats::na.omit()
  
  # Combine everything
  mermaid_code <- paste(
    "```{mermaid}",
    "flowchart LR",
    paste(edges, collapse = "\n"),
    "",
    "classDef done fill:#9be9a8,stroke:#2c7a2c,color:#000;",
    "classDef current fill:#fff3b0,stroke:#b58900,color:#000;",
    "",
    paste(class_lines, collapse = "\n"),
    "```",
    sep = "\n"
  )
  
  return(mermaid_code)
}

```


```{r results='asis'}
#| echo: false
cat(make_mermaid_flow(current_level = "C"))
```


## Research Projects

Below you will find brief descriptions of research projects that are in-progress. For each entry, you will find a title, synopsis, status, and links to any additional information available for that project. Please note, we only list projects that the Research Ring is leading or majorly contributing to. We do not track projects team members collaborate on with other groups.

------------------------------------------------------------------------

### "Deaths of Despair among Protective Service Occupations"

------------------------------------------------------------------------

#### Synopsis:

This project will assess the burden of deaths of despair across major protective service occupations (e.g., police, corrections, fire, EMS/paramedics) using large-scale administrative data.

#### Status:

```{r results='asis'}
#| echo: false
cat(make_mermaid_flow(current_level = "B"))
```

------------------------------------------------------------------------

### "Box breathing and prolonged exhalation reduces markers of physiological stress reactivity in response to a virtual trier social stress test" 

------------------------------------------------------------------------

#### Synopsis:

This randomized controlled trial examined the effects of two brief slow-breathing techniques on physiological and subjective stress responses in a sample of 66 participants exposed to a virtual Trier Social Stress Test (TSST). Compared to a normal breathing control group, both breathing interventions attenuated post-stressor increases in heart rate, state anxiety, and salivary alpha-amylase, indicating a buffering effect on acute sympathetic and psychological stress responses. These findings support the practical utility of ultra-short breathing interventions for stress mitigation, with potential relevance for high-stress occupations such as law enforcement, firefighting, and military service.

#### Status:

```{r results='asis'}
#| echo: false
cat(make_mermaid_flow(current_level = "E"))
```

------------------------------------------------------------------------

### "Moving Targets: Classification Drift and the Changing Profile of Active Shooter Events in the United States" 

------------------------------------------------------------------------

#### Synopsis:

This project will examine longitudinal trends in active shooter incident resolutions using FBI data spanning 2000–2024, with a particular focus on whether the proportion of incidents ending with the attacker fleeing the scene has increased over time. Findings from this project will have direct implications for law enforcement training and operational preparedness, particularly as a growing share of active shooter responses may involve scene stabilization and suspect search operations rather than direct engagement. 

#### Status:

```{r results='asis'}
#| echo: false
cat(make_mermaid_flow(current_level = "C"))
```

------------------------------------------------------------------------

### "Examining Stress Habituation Across Repeated VR Scenario Exposures" 

------------------------------------------------------------------------

#### Synopsis:

This project will examine the effects of repeated exposure to stressful virtual reality (VR) scenarios on acute physiological stress responses, with the aim of evaluating whether habituation occurs across multiple exposures. Participants will complete three separate scenario days using a VR system, with salivary biomarkers collected at each session to index acute stress reactivity. This study is being conducted in collaboration with the MAP Lab at Texas State University.

#### Status:

```{r results='asis'}
#| echo: false
cat(make_mermaid_flow(current_level = "B"))
```

------------------------------------------------------------------------

### "Legal gun ownership and homicides in Brazil" 

------------------------------------------------------------------------

#### Synopsis:

This study uses Bayesian multilevel models to analyze the association between 
legal gun ownership and homicide rates in Brazil. It used official data on 
firearm possession and licenses to carry rather than proxy measures. The 
results suggest that legal gun ownership is unlikely to be a significant driver 
of homicide rates. Preprint available on [CrimRxiv](https://doi.org/10.21428/cb6ab371.b1300c2c){target="_blank"}. 

#### Status:

```{r results='asis'}
#| echo: false
cat(make_mermaid_flow(current_level = "D"))
```

------------------------------------------------------------------------

### "Calibration & validation of a police UoF assessment instrument" 

------------------------------------------------------------------------

#### Synopsis:
This project draws on data from the final evaluations of a defensive tactics class involving over 2,000 Brazilian police recruits. The evaluation used a standard reality-based simulation of a domestic call and an instrument designed to assess police performance in use-of-force incidents. This study now explores, validates, and calibrates that instrument.

#### Status:

```{r results='asis'}
#| echo: false
cat(make_mermaid_flow(current_level = "B"))
```

------------------------------------------------------------------------

### "What officers should be trained for? (Brazil)" 

------------------------------------------------------------------------

#### Synopsis:

This project is a collaboration between U.S.-based scholars and Brazilian pracademics. It uses a mixed-methods approach, combining a national-level survey with over 20,000 responses, data from police reports from a Brazilian police department, and a focus group. Together, these sources are used to identify the use-of-force situations most commonly experienced by officers, the contextual factors that increase risk in those incidents, and other critical situations—all with the goal of informing police use-of-force training. Results will be compared with an ongoing parallel study being conducted with U.S. law enforcement.

#### Status:

```{r results='asis'}
#| echo: false
cat(make_mermaid_flow(current_level = "C"))
```

